---
title: 'Lab 4 - Part 2: Time Series Data'
author: "Indigo Bannister"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)

```

## Read in data

```{r}

toolik <- read_csv(here("data",
                        "toolikweather.csv"))

```

## Convert this into a tsibble

```{r}

toolik_ts <- toolik %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  as_tsibble(key = NULL, index = date)

```

## Visualize

```{r}

ggplot(data = toolik_ts, aes(x = date, y = mean_airtemp)) +
  geom_line()

```

## Use index_by() function to aggregate time series info by intervals we specify 

```{r}

toolik_month <- toolik_ts %>% 
  index_by(yr_mo = ~yearmonth(.)) %>% 
  summarise(monthly_mean_temp = mean(mean_airtemp, na.rm = TRUE))

ggplot(data = toolik_month, aes(x = yr_mo, y = monthly_mean_temp)) +
  geom_line() +
  facet_wrap(~month(yr_mo, label = TRUE))


```
## Aggregate by week

```{r}
toolik_weekly <- toolik_ts %>% 
  index_by(weekly = ~yearweek(.)) %>% 
  summarize(weekly_airtemp = mean(mean_airtemp, na.rm = TRUE),
            min_airtemp = min(mean_airtemp))

head(toolik_weekly)
```

## Filtering Time Series
Use filter_index() to specify ranges of time series to keep or exclude

```{r}

toolik_ts %>% 
  filter_index("2000-06" ~ "2001-10")

toolik_ts %>% 
  filter_index("2006-04-10" ~ "2006-05-15")

```

## Visualizing time series data

### Season plots

```{r}

toolik_ts %>% 
  filter(year(date) > 2014) %>%  #equivalent to using filter_index("2015" ~ .)
  gg_season(y = mean_airtemp)

toolik_month %>% 
  gg_season(y = monthly_mean_temp) #ggseason is a wrapper for ggplot, so can use all of the ggplot customizations

```

### Seasonal subseries plot\

```{r}

toolik_month %>% 
  gg_subseries(monthly_mean_temp)

```

## Find moving average

Using the slider package in R

```{r}

set.seed(2021)
test <- rnorm(n = 100, mean = 40, sd = 10)

slide(test, ~.x, .before = 2, .after = 2)
```


```{r}

roll_toolik_15 <- toolik_ts %>% 
  mutate(ma_15d = as.numeric(slide(toolik_ts$mean_airtemp,
                                   mean,
                                   .before = 7,
                                   .after = 7)))

ggplot() +
  geom_line(data = toolik_ts, 
            aes(x = date, y = mean_airtemp), 
            size = 0.2, 
            color = "gray") + 
  geom_line(data = roll_toolik_15,
            aes(x = date, y = ma_15d),
            color = "red") +
  theme_minimal()

```

## Create the ACF

```{r}

toolik_ts %>% 
  ACF(mean_airtemp) %>% 
  autoplot()

# when correlated by one day, then data is highly correlated

toolik_month %>% 
  ACF(monthly_mean_temp) %>% 
  autoplot()

# when use monthly ACF, see clear seasonality, lag in month
# observations that are 12 months apart are strongly positively correlated. observations 6 months apart are strongly negatively correlated. 

```

## Decomposition by STL

```{r}

toolik_dec <- toolik_month %>% 
  model(STL(monthly_mean_temp ~ season(window = Inf)))

components(toolik_dec) %>% 
  autoplot()

```

